Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-25 arxiv/diff_cov.tex
fix links
wip

\begin{theorem}\label{thm_main_RMT}
	Suppose that $X_1$ and $X_2$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	With high probability over the randomness of $X_1$ and $X_2$, we have the following limits:
	\begin{enumerate}
		\item[i)]  The variance equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following two equations:
			\begin{align}
				a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \bigbrace{\frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{\rho_1}{\rho_1 + \rho_2}. \label{eq_a12extra}
			\end{align}
		\item[ii)] For any fixed vector $w\in\real^p$ with unit norm, the bias equation converges to the following asymptotic limit
			\begin{align}\label{def Pibias}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following two equations with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
				= b_0.
			\end{align}
	\end{enumerate}
\end{theorem}
\noindent{\textbf{Remark.}} Our result extends a well-known result on the inverse of sample covariance matrices.
In particular, by setting $X_1 = 0$, we obtain the second result in Fact \ref{lem_minv}.
When $X_1=0$, it is not hard to solve $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ from equation \eqref{eq_a12extra}, and our result reduces to the case of a single sample covariance matrix.
For the single-task case of a multivariate Gaussian random matrix, the result follows from the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For a non-Gaussian random matrix, the result can be obtained using the Stieltjes transform method (cf. Lemma 3.11 of \citet{bai2009spectral}).

By combining the bias and variance asymptotics, we obtain the generalization error of $\hat{\beta}_2^{\MTL}$.
Since the proof is similar to the same covariates setting, we describe the key differences and omit the rest of the details.
First, we will rescale $X_1$ by $A_1 / A_2$ before appling Theorem \ref{thm_main_RMT}.
Second, we approximate ${X^{(1)}}^{\top}X^{(1)}$ with $n_1 \Sigma^{(1)}$ in equation \eqref{eq_diff_cov} using Fact \ref{lem_minv}.
For the rest of this section, we the main ideas in the proof of Theorem \ref{thm_main_RMT}.
%We will give a simple (although not totally rigorous) derivation of the equations \eqref{eq_a12extra} and \eqref{eq_a34extra}.
The full proof can be found in Section \ref{sec_maintools}.

@@

%For our purpose, we use a convenient linearization trick in linear algebra, that is, the SVD of a rectangular matrix $A$ is equivalent to the study of the eigendecomposition of the symmetric block matrix
%$$H(A):=\begin{pmatrix}0 & A \\ A^\top & 0\end{pmatrix},$$
%which is linear in $A$. This trick has been used in many random matrix literatures, such as \cite{Anisotropic, AEK_Gram, XYY_circular,DY20201}.
For the matrix $Y$ that we want to study, we can also write $Y = A A^{\top}$, where $A = [\Lambda U^\top (Z^{(1)})^\top,V^\top (Z^{(2)})^\top] \in \real^{p\times {n_1 + n_2}}$.
%}\HZ{what does this trick mean? use less technical words},
%This idea dates back at least to Girko, see e.g., the works \cite{girko1975random,girko1985spectral} and references therein.
Consider the following $(p+n)\times (p+n)$ symmetric block matrix
%which is a linear function of $(Z^{(1)})$ and $(Z^{(2)})$:
%\begin{definition}[Linearizing block matrix]\label{def_linearHG}%definiton of the Green function
%We define the $(n+N)\times (n+N)$ block matrix
 \begin{equation}\label{linearize_block}
    H \define (n_1 + n_2)^{-1/2}\left( {\begin{array}{*{20}c}
   { 0 } & \Lambda U^{\top}(Z^{(1)})^\top & V^\top (Z^{(2)})^\top  \\
   {Z^{(1)} U\Lambda  } & {0} & 0 \\
   {Z^{(2)}V} & 0 & 0
   \end{array}} \right).
 \end{equation}
The resolvent of $H$ is defined as follows.
\begin{definition}[Resolvents]
We define the resolvent or Green's function of $H$ as
 $G(z) \define (H - z\id_{p + n_1 + n_2})^{-1}$, for any complex value $z\in \mathbb C$.  %\left[H -\left( {\begin{array}{*{20}c}
%   { z\id_{p}} & 0 & 0 \\
%   0 & { \id_{n_1}}  & 0\\
%      0 & 0  & { \id_{n_2}}\\
%\end{array}} \right)\right]^{-1} , \quad z\in \mathbb C ,
%It is easy to verify that the eigenvalues $\lambda_1(H)\ge \ldots \ge \lambda_{n+N}(H)$ of $H$ are related to the ones of $\mathcal Q_1$ through
%\begin{equation}\label{Heigen}
%\lambda_i(H)=-\lambda_{n+N-i+1}(H)=\sqrt{\lambda_i\left(\mathcal Q_2\right)}, \ \ 1\le i \le n\wedge N, \quad \text{and}\quad \lambda_i(H)=0, \ \ n\wedge N + 1 \le i \le n\vee N.
%\end{equation}
%and
%$$\lambda_i(H)=0, \ \ n\wedge N + 1 \le i \le n\vee N.$$
%where we used the notations $n\wedge N:=\min\{N,M\}$ and $n\vee N:=\max\{N,M\}$.
%\begin{definition}[Index sets]\label{def_index}
Similarly, the resolvent of $Y$ is defined as $\cal G(z):=(Y - z\id_{p})^{-1}$, for any $z\in \mathbb C$.
\end{definition}

Using the Schur complement formula, we have that %\HZ{Has $G(z)$ been defined before?}
 \begin{equation} \label{green2}
  G(z) =  \left( {\begin{array}{*{20}c}
   { \cal G (z)} & \cal G(z)A  \\
   A^\top \cal G(z) & z (A^\top A - z)^{-1}
\end{array}} \right)^{-1}, %\quad \cal G_R:=(W^\top W - z)^{-1} ,
 \end{equation}
This shows that a control of $G(z)$ yields directly a control of $\mathcal G(z)$ as the upper-left block. On the other hand, $G(z)$ is a little  easier to use than $\cal G(z)$, and obviously contains more information.

\noindent\textbf{Asymptotic limits of resolvents.}
We now describe the asymptotic limit of $G(z)$ as $n\to \infty$. First we define the deterministic limits of $(m_1(z), m_{2}(z))$ by $\left(-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z),-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\right)$, where $(a_1(z), a_2(z))$ is
%\HZ{This $M_{1}, M_{2}$ notation is awkward; consider changing it to ${M}_1(z), {M}_2(z)$ (or  better, say $a_1(z), a_2(z)$? if they correspond to $a_1, a_2$ when $z=0$)},
the unique solution to the following system of equations
\begin{equation}\label{selfomega_a}
\begin{split}
& \frac{\rho_1}{a_{1}(z)} = \frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2}{ - z+\lambda_i^2 a_{1}(z) +a_{2} (z) } + (\rho_1+\rho_2),\  \frac{\rho_2}{a_{2}(z)} = \frac{1}{p}\sum_{i=1}^p \frac{1 }{  -z+\lambda_i^2 a_{1}(z) +  a_{2}(z)  }+ (\rho_1+\rho_2) .
\end{split}
\ee
%satisfying that $\im a_{1}(z)< 0$ and $\im a_{2}(z)<0$ for $z\in \C_+$ with $\im z$.
The existence and uniqueness of the solution $(a_1(z), a_2(z))$ for $z$ around 0 will be proved in Section \ref{sec contract}.
%Here, for simplicity of notations, we introduced the following ratios
%\be\label{ratios}
% \gamma_n :=\frac{p}{n}=\frac{1}{\rho_1+\rho_2},\quad r_1 :=\frac{n_1}{n}=\frac{\rho_1}{\rho_1+\rho_2},\quad r_2 :=\frac{n_2}{n}=\frac{\rho_2}{\rho_1+\rho_2}.
%\ee
We define the matrix limit of $G(z)$ as
\be \label{defn_piw}
\Gi(z) := \begin{pmatrix} (-z\id_p+a_{1}(z)\Lambda^2  +  a_{2}(z)\id_p)^{-1} & 0 & 0 \\ 0 & - \frac{\rho_1+\rho_2}{\rho_1} a_{1}(z)\id_{n_1} & 0 \\ 0 & 0 & -\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\id_{n_2}  \end{pmatrix}.\ee
In particular, the matrix limit of $\cal G(z)$ is given by
\be\label{matrix limit}(-z\id_p+a_{1}(z)\Lambda^2 + a_{2}(z)\id_p)^{-1}.\ee

@@

>>>>

\begin{theorem}\label{thm_main_RMT}
	Suppose that $X_1$ and $X_2$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	With high probability over the randomness of $X_1$ and $X_2$, we have the following limits:
	\begin{enumerate}
		\item[i)]  The variance equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following two equations:
			\begin{align}
				a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \bigbrace{\frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{\rho_1}{\rho_1 + \rho_2}. \label{eq_a12extra}
			\end{align}
		\item[ii)] For any fixed vector $w\in\real^p$ with unit norm, the bias equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following two equations with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
				= b_0.
			\end{align}
	\end{enumerate}
\end{theorem}
\noindent{\textbf{Remark.}} Our result extends a well-known result on the inverse of sample covariance matrices.
In particular, by setting $X_1 = 0$, we obtain the second result in Fact \ref{lem_minv}.
When $X_1=0$, it is not hard to solve $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ from equation \eqref{eq_a12extra}, and our result reduces to the case of a single sample covariance matrix.
For the single-task case of a multivariate Gaussian random matrix, the result follows from the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For a non-Gaussian random matrix, the result can be obtained using the Stieltjes transform method (cf. Lemma 3.11 of \citet{bai2009spectral}).

By combining the bias and variance asymptotics, we obtain the generalization error of $\hat{\beta}_2^{\MTL}$.
Since the proof is similar to the same covariates setting, we describe the key differences and omit the rest of the details.
First, we will rescale $X_1$ by $A_1 / A_2$ before appling Theorem \ref{thm_main_RMT}.
Second, we approximate ${X^{(1)}}^{\top}X^{(1)}$ with $n_1 \Sigma^{(1)}$ in equation \eqref{eq_diff_cov} using Fact \ref{lem_minv}.
For the rest of this section, we present the main ideas in the proof of Theorem \ref{thm_main_RMT}.
%We will give a simple (although not totally rigorous) derivation of the equations \eqref{eq_a12extra} and \eqref{eq_a34extra}.
The full proof can be found in Section \ref{sec_maintools}.

@@

%For our purpose, we use a convenient linearization trick in linear algebra, that is, the SVD of a rectangular matrix $A$ is equivalent to the study of the eigendecomposition of the symmetric block matrix
%$$H(A):=\begin{pmatrix}0 & A \\ A^\top & 0\end{pmatrix},$$
%which is linear in $A$. This trick has been used in many random matrix literatures, such as \cite{Anisotropic, AEK_Gram, XYY_circular,DY20201}.
For the matrix $Y$ that we want to study, we can also write $Y = A A^{\top}$, where
	\[ A = (n_1+ n_2)^{-1/2} [\Lambda U^\top (Z^{(1)})^\top,V^\top (Z^{(2)})^\top] \]
is a $p$ by $n_1 + n_2$ matrix.
%}\HZ{what does this trick mean? use less technical words},
%This idea dates back at least to Girko, see e.g., the works \cite{girko1975random,girko1985spectral} and references therein.
Consider the following symmetric and square matrix whose dimension is $p + n_1 + n_2$.
%which is a linear function of $(Z^{(1)})$ and $(Z^{(2)})$:
%\begin{definition}[Linearizing block matrix]\label{def_linearHG}%definiton of the Green function
%We define the $(n+N)\times (n+N)$ block matrix
 \begin{equation}\label{linearize_block}
    H \define (n_1 + n_2)^{-1/2}\left( {\begin{array}{*{20}c}
   { 0 } & \Lambda U^{\top}(Z^{(1)})^\top & V^\top (Z^{(2)})^\top  \\
   {Z^{(1)} U\Lambda  } & {0} & 0 \\
   {Z^{(2)}V} & 0 & 0
   \end{array}} \right).
 \end{equation}
The resolvent of $H$ is defined as follows.
\begin{definition}[Resolvents]
The resolvent or Green's function of $H$ is given by
	\begin{align*}
    G(z) \define (H - z\id)^{-1}, \text{ for any complex value } z\in \mathbb C.  %\left[H -\left( {\begin{array}{*{20}c}
	\end{align*}
%   { z\id_{p}} & 0 & 0 \\
%   0 & { \id_{n_1}}  & 0\\
%      0 & 0  & { \id_{n_2}}\\
%\end{array}} \right)\right]^{-1} , \quad z\in \mathbb C ,
%It is easy to verify that the eigenvalues $\lambda_1(H)\ge \ldots \ge \lambda_{n+N}(H)$ of $H$ are related to the ones of $\mathcal Q_1$ through
%\begin{equation}\label{Heigen}
%\lambda_i(H)=-\lambda_{n+N-i+1}(H)=\sqrt{\lambda_i\left(\mathcal Q_2\right)}, \ \ 1\le i \le n\wedge N, \quad \text{and}\quad \lambda_i(H)=0, \ \ n\wedge N + 1 \le i \le n\vee N.
%\end{equation}
%and
%$$\lambda_i(H)=0, \ \ n\wedge N + 1 \le i \le n\vee N.$$
%where we used the notations $n\wedge N:=\min\{N,M\}$ and $n\vee N:=\max\{N,M\}$.
%\begin{definition}[Index sets]\label{def_index}
Likewise, the resolvent of $Y$ is given by $\cal G(z) \define (Y - z\id)^{-1}$, for any $z\in \mathbb C$.
\end{definition}
Let $m_1(z):= \frac{1}{n_1}\sum_{i = p+1}^{p + n_1} G_{i, i}(z)$ and $m_2(z):= \frac{1}{n_2}\sum_{i = p + n_1 + 1}^{p + n_1 + n_2} G_{i, i}(z)$.
Using the Schur complement formula, we have that %\HZ{Has $G(z)$ been defined before?}
 \begin{equation} \label{green2}
  G(z) =  \left( {\begin{array}{*{20}c}
   { \cal G (z)} & \cal G(z)A  \\
   A^\top \cal G(z) & z (A^\top A - z)^{-1}
\end{array}} \right)^{-1}, %\quad \cal G_R:=(W^\top W - z)^{-1} ,
 \end{equation}
This shows that a control of $G(z)$ yields directly a control of $\mathcal G(z)$ as the upper-left block. On the other hand, $G(z)$ is a little  easier to use than $\cal G(z)$, and obviously contains more information.

%\noindent\textbf{Asymptotic limits of resolvents.}
We now describe the asymptotic limit of $G(z)$ as $p$ goes to infinity.
First, we define the deterministic limits of $(m_1(z), m_{2}(z))$ by $\left(-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z),-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\right)$, where $(a_1(z), a_2(z))$ is
%\HZ{This $M_{1}, M_{2}$ notation is awkward; consider changing it to ${M}_1(z), {M}_2(z)$ (or  better, say $a_1(z), a_2(z)$? if they correspond to $a_1, a_2$ when $z=0$)},
the unique solution to the following system of equations
\begin{equation}\label{selfomega_a}
\begin{split}
& \frac{\rho_1}{a_{1}(z)} = \frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2}{ - z+\lambda_i^2 a_{1}(z) +a_{2} (z) } + (\rho_1+\rho_2),\  \frac{\rho_2}{a_{2}(z)} = \frac{1}{p}\sum_{i=1}^p \frac{1 }{  -z+\lambda_i^2 a_{1}(z) +  a_{2}(z)  }+ (\rho_1+\rho_2) .
\end{split}
\ee
%satisfying that $\im a_{1}(z)< 0$ and $\im a_{2}(z)<0$ for $z\in \C_+$ with $\im z$.
The existence and uniqueness of the solution $(a_1(z), a_2(z))$ for $z$ around 0 will be proved in Section \ref{sec contract}.
%Here, for simplicity of notations, we introduced the following ratios
%\be\label{ratios}
% \gamma_n :=\frac{p}{n}=\frac{1}{\rho_1+\rho_2},\quad r_1 :=\frac{n_1}{n}=\frac{\rho_1}{\rho_1+\rho_2},\quad r_2 :=\frac{n_2}{n}=\frac{\rho_2}{\rho_1+\rho_2}.
%\ee
We define the matrix limit of $G(z)$ as
\be \label{defn_piw}
\Gi(z) := \begin{pmatrix} (-z\id_p+a_{1}(z)\Lambda^2  +  a_{2}(z)\id_p)^{-1} & 0 & 0 \\ 0 & - \frac{\rho_1+\rho_2}{\rho_1} a_{1}(z)\id_{n_1} & 0 \\ 0 & 0 & -\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\id_{n_2}  \end{pmatrix}.\ee
In particular, the matrix limit of $\cal G(z)$ is given by
\be\label{matrix limit}(-z\id_p+a_{1}(z)\Lambda^2 + a_{2}(z)\id_p)^{-1}.\ee

@@

\paragraph{Notations.}
Define the index sets
$$\cal I_0:=\llbracket 1,p\rrbracket, \quad  \cal I_1:=\llbracket p+1,p+n_1\rrbracket, \quad \cal I_2:=\llbracket p+n_1+1,p+n_1+n_2\rrbracket ,\quad \cal I:=\cal I_0\cup \cal I_1\cup \cal I_2  .$$
 We will consistently use the latin letters $i,j\in\sI_{0}$ and greek letters $\mu,\nu\in\sI_{1}\cup \sI_{2}$. Correspondingly, the indices of the matrices $Z^{(1)}$ and $Z^{(2)}$ are labelled as
 \be\label{labelZ}
 Z^{(1)}= (Z^{(1)}_{\mu i}:i\in \mathcal I_0, \mu \in \mathcal I_1), \quad Z^{(2)}= (Z^{(2)}_{\nu i}:i\in \mathcal I_0, \nu \in \mathcal I_2).\ee

Moreover, we also define the following averaged resolvents, which are the (weighted) partial traces of $G$:
\be\label{defm}
\begin{split}
m(z) :=\frac1p\sum_{i\in \cal I_0} G_{ii}(z) ,\quad & m_0(z):=\frac1p\sum_{i\in \cal I_0} \lambda_i^2 G_{ii}(z),\\
 m_1(z):= \frac{1}{n_1}\sum_{\mu \in \cal I_1}G_{\mu\mu}(z) ,\quad & m_2(z):= \frac{1}{n_2}\sum_{\nu\in \cal I_2}G_{\nu\nu}(z) .
\end{split}
\ee
We will show that these averaged resolvents satisfy some deterministic self-consistent equations asymptotically, which will be the core part of the proof. We refer the reader to the discussions below \eqref{0self_Gii} and \eqref{0self_Gmu1}.

\subsection{Warmup: The Gaussian Case}
Now we give a heuristic derivation of the matrix limit when the entries of $Z^{(1)}$ and $Z^{(2)}$ are i.i.d. Gaussian. In this case,
%However, notice that if the entries of $(Z^{(1)})\equiv (Z^{(1)})^{\text{Gauss}}$ and $(Z^{(2)})\equiv (Z^{(2)})^{\text{Gauss}}$ are i.i.d. Gaussian, then
by the rotational invariance of the multivariate Gaussian distribution we have
\be\label{eq in Gauss} Z^{(1)} U\Lambda \stackrel{d}{=} Z^{(1)} \Lambda, \quad Z^{(2)} V \stackrel{d}{=} Z^{(2)},\ee
where ``$\stackrel{d}{=}$" means ``equal in distribution". Hence it suffices to consider the following resolvent
 \begin{equation} \label{resolv Gauss1}
   G(z)= \left( {\begin{array}{*{20}c}
   { -z\id_{p} } & n^{-1/2}\Lambda (Z^{(1)})^\top & n^{-1/2} (Z^{(2)})^\top  \\
   {n^{-1/2} Z^{(1)} \Lambda  } & {-\id_{n_1}} & 0 \\
   {n^{-1/2} Z^{(2)}} & 0 & {-\id_{n_2}}
   \end{array}} \right)^{-1}.
 \end{equation}

\medskip
\noindent{\bf Schur complements.}
%Now we discuss about how to obtain the matrix limit in equation \eqref{defn_piw}.
%\HZ{Divide into several parts like sec 7 to improve readability.}
The core quantities of the derivation are the following resolvent minors, which are defined by removing certain rows and columns of the matrix $H$.
\begin{definition}[Resolvent minors]\label{defn_Minor}
 For any $ (p+n)\times (p+n)$ matrix $\cal A$ and $a\in \mathcal I$, we define the minor $\cal A^{(a)}:=(\cal A_{a_1 a_2}:a_1, a_2 \in \mathcal I\setminus \{a\})$ as the $ (p+n-1)\times (p+n-1)$ matrix obtained by removing the $a$-th row and column in $\cal A$. Note that we keep the names of indices when defining $\cal A^{(a)}$, i.e. $(\cal A^{(a)})_{a_1a_2}= \cal A_{a_1 a_2}$ for $a_1,a_2\ne a$. Correspondingly, we define the resolvent minor for $G$ in equation \eqref{resolv Gauss1} as %(recall equation \eqref{green2})
\begin{align*}
G^{(\mathfrak c)}:&=\left[ \left( {\begin{array}{*{20}c}
   { -z\id_{p} } & n^{-1/2}\Lambda (Z^{(1)})^\top & n^{-1/2} (Z^{(2)})^\top  \\
   {n^{-1/2} Z^{(1)} \Lambda  } & {-\id_{n_1}} & 0 \\
   {n^{-1/2} Z^{(2)}} & 0 & {-\id_{n_2}}
   \end{array}} \right)^{(a)}\right]^{-1} ,
%= \left( {\begin{array}{*{20}c}
%   { \mathcal G^{(\mathbb T)}} & \mathcal G^{(\mathbb T)} W^{(\mathbb T)}  \\
%   {\left(W^{(\mathbb T)}\right)^\top\mathcal G^{(\mathbb T)}} & { \mathcal G_R^{(\mathbb T)} }  \\
%\end{array}} \right)  ,
\end{align*}
and define the partial traces $m^{(a)}$, $m_0^{(a)}$, $m_1^{(a)}$ and $m_2^{(a)}$ by replacing $G$ with $G^{(a)}$ in equation \eqref{defm}. We adopt the convention that for the resolvent minor $G^{(a)}$ defined as above, $G^{(a)}_{a_1a_2} = 0$ if $a_1 =a$ or $a_2=a$.
\end{definition}
%Note that the resolvent minor $G^{(\mathfrak c)}$ is defined such that it is independent of the entries in the $\mathfrak c$-th row and column of $H$. One will see a crucial use of this fact in the heuristic proof below.
 Using Schur complement formulas in equation (\ref{resolvent2}), we have that for $i \in \cal I_0$, %$\mu \in \cal I_1$ and $\nu\in \cal I_2$,
%\HZ{This part needs more explanation - cannot understand.}
\begin{align}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu,\nu\in \mathcal I_1} Z^{(1)}_{\mu i}Z^{(1)}_{\nu i}G^{\left( i \right)}_{\mu\nu} - \frac{1}{n} \sum_{\mu,\nu\in \mathcal I_2} Z^{(2)}_{\mu i}Z^{(2)}_{\nu i}G^{\left( i \right)}_{\mu\nu} -2 \frac{\lambda_i}{n} \sum_{\mu\in \cal I_1,\nu\in \mathcal I_2} Z^{(1)}_{\mu i}Z^{(2)}_{\nu i}G^{\left( i \right)}_{\mu\nu} , \label{0self_Gii}
\end{align}
and for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\begin{align}
\frac{1}{{G_{\mu\mu} }}&=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}\lambda_i \lambda_j Z^{(1)}_{\mu i}Z^{(1)}_{\mu j} G^{\left(\mu\right)}_{ij}, \quad \frac{1}{{G_{\nu\nu} }}=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}  Z^{(2)}_{\nu i}Z^{(2)}_{\nu j}  G^{\left(\nu\right)}_{ij},
\label{0self_Gmu1}
\end{align}
where we recall the notations in equation \eqref{labelZ}. To see why equation \eqref{0self_Gii} holds, we have that by Schur complement formula,
\begin{align*}
\frac{1}{G_{ii}}= -z - H_i G^{(i)}H_{i}^\top, \quad H_i = \left( \mathbf 0_{p-1}, (n^{-1/2}\lambda_i (Z^{(1)})^\top_{i\mu }:\mu \in \cal I_1),(n^{-1/2} (Z^{(2)})^\top_{i\nu }:\nu \in \cal I_2)\right),
\end{align*}
where $H_i$ is actually the $i$-th row of $H$ with the $(i,i)$-th entry removed. Expanding the above expression, we obtain equation \eqref{0self_Gii}. The two expressions in equation \eqref{0self_Gmu1} are easier to obtain using Schur complement formula.

\medskip
\noindent{\bf Concentration estimates.} Now for the right-hand side of equation \eqref{0self_Gii}, notice that the resolvent minor $G^{(i)}$ is defined such that it is independent of the entries $Z^{(1)}_{\mu i}$ and $Z^{(2)}_{\nu i}$. Hence by the concentration inequalities in Lemma \ref{largedeviation}, we have that the  right-hand side of equation \eqref{0self_Gii} concentrates around the partial expectation over the entries $\{Z^{(1)}_{\mu i}: \mu \in \cal I_1 \}\cup \{Z^{(2)}_{\nu i}: \nu \in \cal I_2\}$, i.e., with overwhelming probability,
\begin{align*}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu \in \mathcal I_1}  G^{\left( i \right)}_{\mu\mu} - \frac{1}{n} \sum_{\mu\in \mathcal I_2} G^{\left( i \right)}_{\mu\mu} +\oo(1)= - z - \lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1^{(i)}(z)-  \frac{\rho_2}{\rho_1+\rho_2} m_2^{(i)}(z)+\oo(1),
\end{align*}
where we used the definition of $m_1^{(i)}$ and $m_2^{(i)}$ in equation \eqref{defm} with $G$ replaced by $G^{(i)}$. Intuitively, since we have removed only one column and one row out of the $(p+n)$ columns and rows in $H$, $m_1^{(i)}$ and $m_2^{(i)}$ should be close to the original $m_1$ and $m_2$. Hence we obtain from the above equation that
\begin{align}\label{1self_Gii}
 G_{ii}  = -\left( z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1 +  \frac{\rho_2}{\rho_1+\rho_2}m_2+\oo(1)\right)^{-1}.
\end{align}
Similarly, we can obtain from equation \eqref{0self_Gmu1} that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{1self_Gmu} G_{\mu \mu }=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},\ee
with overwhelming probability. The rigorous derivation of the above concentration estimates will be given in the proof of Lemma \ref{lemm_selfcons_weak}; see equations \eqref{self_Gii}-\eqref{erri}.

\medskip
\noindent\textbf{Deriving the self-consistent equations.} Now taking average of equation \eqref{1self_Gmu}, we obtain that
\be\label{2self_Gmu} m_1= \frac{1}{n_1}\sum_{\mu \in \cal I_1}G_{\mu\mu}=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad m_2=\frac{1}{n_2}\sum_{\nu \in \cal I_2}G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},
\ee
with overwhelming probability. As a byproduct, comparing equations \eqref{1self_Gmu} and \eqref{2self_Gmu}, we obtain that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{2.5self_Gmu} G_{\mu \mu }=m_1 +\oo(1),\quad G_{\nu\nu}=m_2+\oo(1),\ee
with overwhelming probability. Together with the definition of $m$ and $m_0$ in equation \eqref{defm}, the two equations in equation \eqref{2self_Gmu} give that
\be\label{3self_Gmu}  \frac1{m_1}= -1- \frac{1}{n} \sum_{i=1}^p \lambda_i^2 G_{ii}+ \oo(1),\quad \frac1{m_2}=-1-\frac{1}{n} \sum_{i=1}^p G_{ii}  + \oo(1),\ee
with overwhelming probability. Plugging equation \eqref{1self_Gii} into equation \eqref{3self_Gmu}, we obtain that
\be\label{approximate m1m2}
\begin{split}
& \frac1{m_1}= -1+ \frac{1}{n} \sum_{i=1}^p \frac{\lambda_i^2 }{ z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)}+ \oo(1),\\
& \frac1{m_2}=-1+\frac{1}{n} \sum_{i=1}^p \frac{1}{ z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)}  + \oo(1),
\end{split}
\ee
with overwhelming probability, which give a system of approximate self-consistent equations for $(m_1,m_2)$.
A rigorous derivation of these two equations will be given in the proof of Lemma \ref{lemm_selfcons_weak}. Compare \eqref{approximate m1m2} to the deterministic self-consistent equations in equation \eqref{selfomega_a}, one can observe that we should have
\be\label{approx m12 add}
(m_1,m_2) =\left(-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z),-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\right)+\oo(1) \quad \text{with overwhelming probability. }
\ee
To justify this identity rigorously, we need to know that the self-consistent equations are stable, that is, a small perturbation of the equations leads to a small perturbation of the solution. This will be given by Lemma \ref{lem_stabw}.
